---
title: "Data 606 Final Project Proposal"
author: "Mohamed Hassan-El Serafi"
date: "2023-04-09"
output: html_document
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
library(tidyverse)
library(DT)
library(psych)
df <- read.csv("https://raw.githubusercontent.com/moham6839/Data606_ProjectProposal/main/nbaplayersdraft.csv")
df
```


```{r}
sum(is.na(df))
```

```{r}
summary(df)
```



```{r}
glimpse(df)
```



```{r}
# Drop Rank column, since it is the same as overall_pick
df <- select(df, -rank)
head(df)
```


```{r}
new_df <- df %>%
  select(overall_pick, win_shares, minutes_played, games, years_active, box_plus_minus)
head(new_df)
```



### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

What factors affect a player's WinShares? 



### Cases 

**What are the cases, and how many are there?**

There are 1,922 cases, with 24 columns. Each case is a NBA player who was drafted between 1989 and 2021.

### Data collection 

**Describe the method of data collection.**

According to the [Kaggle](https://www.kaggle.com/datasets/mattop/nba-draft-basketball-player-data-19892021) link where the dataset was downloaded from, the data was collected from Basketball Reference, which you can find [here](https://www.basketball-reference.com/draft/).

### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

I found the dataset on Kaggle. Here is the link: https://www.kaggle.com/datasets/mattop/nba-draft-basketball-player-data-19892021


### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable will be WinShares, which is a quantitative variable.

### Independent Variable(s)

My independent variables will be based on what could affect how high or low WinShares of players, such as overall_pick, games, years_active, minutes_played, minutes_played, and box_plus_minus. Overall draft picks are categorical, while the other variables are numerical.


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**



```{r}
describe(new_df$win_shares)
```


```{r}
describe(new_df$games)
```


```{r}
describe(new_df$years_active)
```


```{r}
describe(new_df$box_plus_minus)
```


```{r}
describe(new_df$overall_pick)
```


```{r}
describe(new_df$minutes_played)
```


```{r}
summary(new_df)
```


```{r}
colSums(is.na(new_df))
```



```{r}
new_df %>% 
  ggplot(aes(x=win_shares)) +
  geom_histogram(bins = 100) +
  labs(title="Amount of WinShares for Each Player",
       x="Number of WinShares",
       y="Count")
```


The histogram shows a skewness to the right for WinShares, with a large amount of player WinShares at or around zero.




```{r}
new_df %>%
  ggplot(aes(box_plus_minus, na.rm=TRUE)) +
  geom_histogram(bins = 50) +
  labs(title="Plus/Minus for Each Player",
       x="Plus/Minus",
       y="Count")
  
```


```{r}
new_df %>%
  ggplot(aes(win_shares, games, na.rm=TRUE)) +
  geom_point() +
  labs(title="Amount of WinShares Based on Games Played",
       x="WinShares",
       y="Games")
```


```{r}
new_df %>%
  ggplot(aes(box_plus_minus, win_shares, na.rm=TRUE)) +
  geom_point() +
  labs(title="Amount of WinShares Based on Plus/Minus of Each Player",
       x="Plus/Minus",
       y="WinShares")
```


```{r}
new_df %>%
  ggplot(aes(overall_pick, win_shares, na.rm=TRUE)) +
  geom_point() +
  labs(title="Amount of WinShares Based on Overall Pick of Players",
       x="WinShares",
       y="Overall Draft Picks")
```

```{r}
new_df %>%
  ggplot(aes(years_active, win_shares, na.rm=TRUE)) +
  geom_point() +
  labs(title="Amount of WinShares Based on Number of Playing Years",
       x="Years Played",
       y="WinShares")
```


```{r}
new_df %>%
  ggplot(aes(minutes_played, win_shares, na.rm=TRUE)) +
  geom_point() +
  labs(title="Amount of WinShares Based on Career Minutes Played",
       x="Career Minutes Played",
       y="WInShares")
```



